---
format: 
  revealjs:
    theme: [style.scss, default]
    width: "100%"
    height: "100%"
    minScale: 1
    maxScale: 1
    center: false
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, error = F)
```

```{r, echo = F}
library(ggplot2)
library(dplyr)
library(cowplot)
library(extrafont)
library(tidyverse)
library(magick)

theme_set(
  theme_bw() + 
    theme(plot.background = element_rect(fill = "#eaebff", color = "#eaebff"),
          legend.box.background = element_rect(color = "black"))
)
```

##

## Basic Biostatistics and Bioinformatics

A seminar series on the fundamentals 

Organised by *SLUBI* and *Statistics at SLU*

Presentation of background and a practical exercise

<br>
Topics

- 3 June. Git and Github
- 17 June. Projects in R
- 1 July. Shiny

<br>
Topic suggestions are welcome

##

**SLUBI**

- SLU bioinformatics center
- Weekly online drop-in (Wednesdays at 13.00)
- slubi@slu.se, [https://www.slubi.se](https://www.slubi.se)
- Alnarp: Lizel Potgieter (Dept. of Plant Breeding)

<br>

**Statistics at SLU**

- SLU statistics center
- Free consultations for all SLU staff
- statistics@slu.se
- Alnarp: Jan-Eric Englund and Adam Flöhr (Dept. of Biosystems and Technology)

## Git and GitHub Presentation

**Introduction**

- **Git** is a version control system for tracking changes in source code during software development.
- **GitHub** is a web-based platform that uses Git for version control and provides collaboration features.

## What is Git?

Definition and Purpose

- **Git** is a distributed version control system.
- It helps developers collaborate on projects.
- Allows tracking changes in source code.

## Key Features of Git

Why Use Git?

- **Branching and Merging**: Create separate branches for new features, merge them back to the main branch.
- **Distributed Development**: Each developer has a complete local copy of the repository.
- **Staging Area**: Review and manage changes before committing them.
- **History and Versioning**: Keep track of every change with commit history.

## Basic Git Commands

Essential Commands to Get Started

- `git init`: Initialize a new Git repository.
- `git clone [url]`: Clone an existing repository.
- `git add [file]`: Stage changes for commit.
- `git commit -m "[message]"`: Commit changes with a message.
- `git status`: Check the status of changes.
- `git log`: View commit history.

## What is GitHub?

Overview and Features

- **GitHub** is a cloud-based platform for version control and collaboration.
- Built on top of Git.
- Provides a web interface for repositories.
- Facilitates collaboration with features like pull requests, issues, and wikis.

## Key Features of GitHub

Why Use GitHub?

- **Repositories**: Store and manage project files.
- **Pull Requests**: Propose and review changes before merging.
- **Issues**: Track bugs and feature requests.
- **GitHub Actions**: Automate workflows with CI/CD.
- **Collaborative Tools**: Wikis, project boards, and discussions.

## Basic GitHub Workflow

How to Collaborate Using GitHub

1. **Fork**: Create a personal copy of someone else's project.
2. **Clone**: Download your fork to your local machine.
3. **Commit**: Make changes and commit them locally.
4. **Push**: Upload changes to your GitHub repository.
5. **Pull Request**: Propose changes to the original repository.
6. **Merge**: Integrate changes after review.

## Using Git and GitHub Together

Integrated Workflow

- **Local Development**: Use Git for version control on your local machine.
- **Remote Collaboration**: Use GitHub to share your repository and collaborate with others.
- **Sync Changes**: Regularly push and pull changes between local and remote repositories.


## Best Practices

Tips for Effective Use

- **Commit Often**: Make frequent commits with meaningful messages.
- **Branching Strategy**: Use branches for features, fixes, and experiments.
- **Code Reviews**: Regularly review code via pull requests.
- **Documentation**: Maintain good documentation for your projects.
- **Backup**: Regularly push changes to remote repositories.


## Resources

Learn More

- **Official Git Documentation**: [git-scm.com/doc](https://git-scm.com/doc)
- **GitHub Learning Lab**: [lab.github.com](https://lab.github.com)
- **Pro Git Book**: [git-scm.com/book/en/v2](https://git-scm.com/book/en/v2)
- **GitHub Guides**: [guides.github.com](https://guides.github.com)

## Q&A

Questions and Answers

- Feel free to ask any questions about Git and GitHub.
- Let's discuss how you can implement these tools in your projects.
